Ch3 process

### 2022-02-18 Ch3_samplesize/Ch3_process.Rmd
Fixed the errors so now all functions work.

lomax.mean <- function(alpha=alpha, k=k){
  return(alpha/(k+1))
}

@@

  nbetasp1      <- length(init.betas)
  k             <- exp(guess[1])
  Xbeta         <- designmat%*%guess[2:nbetasp1]
  alphas        <- exp(Xbeta) # because alpha = ln(X*betas)
  n             <- length(Y)

@@

  mles          <- opt.out$par
  nll.hat       <- opt.out$value
  BIC.mod       <- 2*nll.hat + nbetas*log(length(Y))
  Xbeta.hat     <- designmat%*%mles[-1]
  alphas.hat    <- exp(Xbeta.hat)
  #sumlogapy.hat <- sum(log(alphas.hat+Y))
  k.hat         <-  mles[1] #n/(sumlogapy.hat - sum(Xbeta.hat))

@@

  # The glm.lomax was giving me an error. Kept saying it couldn't find the init betas but worked when I ran it line by line.

  # From the glm framework, we are not including covariates here
  # The next three lines are the arguments to the glm lomax function
  formula=~1
  my.dataf = ith.samples
  response = ith.samples$x

  # This is the glm lomax function deconstructed
  Y           <- response
  n           <- length(Y)
  designmat   <- model.matrix(formula, data=my.dataf)
  nbetas      <- ncol(designmat)
  init.betas  <- c(4,rep(5,nbetas))

  mles          <- opt.out$par
  nll.hat       <- opt.out$value
  BIC.mod       <- 2*nll.hat + nbetas*log(length(Y))
  Xbeta.hat     <- designmat%*%mles[-1]
  alphas.hat    <- exp(Xbeta.hat)
  #sumlogapy.hat <- sum(log(alphas.hat+Y))
  k.hat         <-  mles[1] #n/(sumlogapy.hat - sum(Xbeta.hat))

  mles.sampsizes$alphas[i] <- alphas.hat
  mles.sampsizes$khats[i] <- k.hat
}

mles.sampsizes %>% 
  mutate(mean = lomax.mean(alphas, khats),
         tail100 = lomax.cdf(100, alphas, khats),
         tail250 = lomax.cdf(250, alphas, khats),
         tail500 = lomax.cdf(500, alphas, khats),
         tail750 = lomax.cdf(750, alphas, khats),
         tail1k = lomax.cdf(1000, alphas, khats)) %>% 
  #mutate(across(c(2:8), signif, digits=3)) %>% 
  mutate(across(starts_with("tail"), signif, digits=3))

```
```{r}
# But if we do 1-cdf, which should be P(X>=x) we get high numbers.

1-lomax.cdf(thresh.vals, alpha = mles.sampsizes$alphas[1], k = mles.sampsizes$khats[1])

```
```{r eval=FALSE, echo=FALSE}
###################################
# Run the experiments with different sample sizes and tail thresholds

@@

  formula=~1
  my.dataf = ith.samples
  response = ith.samples$x

  mles          <- opt.out$par
  nll.hat       <- opt.out$value
  BIC.mod       <- 2*nll.hat + nbetas*log(length(Y))
  Xbeta.hat     <- designmat%*%mles[-1]
  alphas.hat    <- exp(Xbeta.hat)
  #sumlogapy.hat <- sum(log(alphas.hat+Y))
  k.hat         <-  mles[1] #n/(sumlogapy.hat - sum(Xbeta.hat))

  mod1 <- list(opt.out = opt.out, designmat=designmat,Y=Y, mles=mles, nll.hat=nll.hat, BIC.mod = BIC.mod,
                   alphas.hat=alphas.hat, k.hat=k.hat,data=my.dataf)

@@

  ith.q <- all.qtests[i]
  ith.cdf <- 1- lomax.cdf(x = ith.q, alpha = ith.a, k = ith.k)
  cdfs.hat[i] <- ith.cdf

@@

```{r results='hide'}
nreps <- 100
gp.mles.reps <- data.frame(NULL)

@@

@@

>>>>

lomax.mean <- function(alpha=alpha, k=k){
  return(alpha/(k-1))
}

@@

  nbetasp1      <- length(guess)
  k             <- exp(guess[1])
  Xbeta         <- designmat%*%guess[2:nbetasp1]
  alphas        <- exp(Xbeta) # because alpha = ln(X*betas)
  n             <- length(Y)

@@

  mles          <- opt.out$par
  nll.hat       <- opt.out$value
  BIC.mod       <- 2*nll.hat + nbetas*log(length(Y))
  Xbeta.hat     <- designmat%*%mles[-1]
  alphas.hat    <- log(Xbeta.hat)
  #sumlogapy.hat <- sum(log(alphas.hat+Y))
  k.hat         <-  mles[1] #n/(sumlogapy.hat - sum(Xbeta.hat))

@@

  mod1 <- lomax.glm(formula = ~1, my.dataf = ith.samples, response = ith.samples$x)
  ith.a <- mod1$alphas.hat[1]
  ith.k <- mod1$k.hat

  mles.sampsizes$alphas[i] <- ith.a
  mles.sampsizes$khats[i] <- ith.k
}

mles.sampsizes %>% 
  mutate(mean = lomax.mean(alphas, khats),
         tail100 = lomax.st(100, alphas, khats),
         tail250 = lomax.st(250, alphas, khats),
         tail500 = lomax.st(500, alphas, khats),
         tail750 = lomax.st(750, alphas, khats),
         tail1k = lomax.st(1000, alphas, khats)) %>% 
  #mutate(across(c(2:8), signif, digits=3)) %>% 
  mutate(across(starts_with("tail"), signif, digits=3))

```{r}
###################################
# Run the experiments with different sample sizes and tail thresholds

@@

  mod1 <- lomax.glm(formula = ~1, my.dataf = ith.samples, response = ith.samples$x)

@@

  ith.q <- all.qtests[i]
  ith.cdf <- lomax.st(x = ith.q, alpha = ith.a, k = ith.k)
  cdfs.hat[i] <- ith.cdf

@@

```{r eval=FALSE, results='hide'}
# I already ran this chunk and it takes a very long time (a night), so saved the result and I'm just loading it in the next chunk
nreps <- 1000
gp.mles.reps <- data.frame(NULL)

@@

```{r}
load("Ch3_samplesize/testdf.RData")
```

@@

Restrict the y axis

gp.mles.reps %>% 
  mutate(tru.tails = rep(tru.cdfs$w.cdfs, n.tru.cdfs),
         tails.ratio = tail/tru.tails,
         thresh = factor(threshold),
         samp.size = factor(samp.size)) %>%
  drop_na(tails.ratio) %>% 
  mutate(nice.tails = signif(tails.ratio, 3)) %>% 
  ggplot(., aes(x = samp.size, y = tails.ratio, color = thresh)) +
  geom_boxplot() +
  labs(title = "Lomax") +
  lims(y = c(-0.1, 3)) +
  theme_minimal() -> lomax.tail

gp.mles.reps %>% 
  mutate(tru.tails = rep(tru.cdfs$w.cdfs, n.tru.cdfs),
         gp.tails.ratio = gp.tail/tru.tails,
         thresh = factor(threshold),
         samp.size = factor(samp.size)) %>%
  mutate(nice.tails = signif(gp.tails.ratio, 3)) %>% 
  ggplot(., aes(x = samp.size, y = gp.tails.ratio, color = thresh)) +
  geom_boxplot() +
  labs(title = "GP") +
  lims(y = c(-0.1, 3)) +
  theme_minimal() -> gp.tail

Check the mean:

```{r}
gp.mles.reps %>% 
  mutate(tru.tails = rep(tru.cdfs$w.cdfs, n.tru.cdfs),
         tails.ratio = tail/tru.tails,
         gp.tails.ratio = gp.tail/tru.tails) %>% 
  group_by(samp.size, threshold) %>% 
  summarise(mean.ratio = mean(tails.ratio),
            mean.gp.ratio = mean(gp.tails.ratio),
            ste.gp.ratio = sd(gp.tails.ratio)/sqrt(length(gp.tails.ratio))) -> tail.ratio.means

tail.ratio.means %>%
  knitr::kable()
```

```{r fig.width=8}
tail.ratio.means %>% 
  ggplot(., aes(x = factor(threshold), y = mean.gp.ratio, color = factor(samp.size))) +
  #geom_point() +
  geom_jitter(size = 2, width = 0.2, alpha = 0.8) +
  #scale_y_log10() +
  theme_minimal()
```
